The goal of {epikit} is to provide miscellaneous functions for 
This is a product of the R4EPIs project; learn more at <https://r4epis.netlify.com>.

## Installation

You can install {epikit} from CRAN:

``` r
install.packages("epikit")
```

<details>
<!--
NOTE: everything inside the details tag will be collapsed and effectively
hidden from the user
 -->
<summary style='text-decoration: underline'>Click here for alternative installation options</summary>
If there is a bugfix or feature that is not yet on CRAN, you can install it via
the {drat} package:


You can install {epikit} from the R4EPI repository:

```{r eval = FALSE}
# install.packages("drat")
drat::addRepo("R4EPI")
install.packages("epikit")
```

You can also install the in-development version from GitHub using the {remotes}
package (but there's no guarantee that it will be stable):

```{r eval = FALSE}
# install.packages("remotes")
remotes::install_github("R4EPI/epikit") 
```

</details>

-------

```{r load_packages}
library("epikit")
```

The {epikit} was primarily designed to house convenience functions for field
epidemiologists to use in tidying their reports. The functions in {epikit} come
in a few categories:

## Give me a break

If you need a quick function to determine the number of breaks you need for a
color scale, you can use `find_breaks()`. This will always start from 1, so that
you can include zero in your scale when you need to.

```{r find_breaks}
find_breaks(100) # four breaks from 1 to 100
find_breaks(100, snap = 20) # four breaks, snap to the nearest 20
find_breaks(100, snap = 20, ceiling = TRUE) # include the highest number
```

## Table modification

These functions all modify the appearance of a table displayed in a report and
work best with the `knitr::kable()` function.

 - `rename_redundant()` renames redundant columns with a single name. (e.g. `hopitalized_percent` and `confirmed_percent` can both be renamed to `%`)
 - `augment_redundant()` is similar to `rename_redundant()`, but it modifies the redundant column names (e.g. `hospitalized_n` and `confirmed_n` can become `hospitalized (n)` and `confirmed (n)`)
 - `merge_ci()` combines estimate, lower bound, and upper bound columns into a single column.

```{r table_mods, results = 'asis', eval = requireNamespace("knitr") && requireNamespace("magrittr")}
library("knitr")
library("magrittr")
df <- data.frame(
  `a n` = 1:6,
  `a prop` = round((1:6) / 6, 2),
  `a deff` = round(pi, 2),
  `b n` = 6:1,
  `b prop` = round((6:1) / 6, 2),
  `b deff` = round(pi * 2, 2),
  check.names = FALSE
)
knitr::kable(df)
df %>%
  rename_redundant("%" = "prop", "Design Effect" = "deff") %>%
  augment_redundant(" (n)" = " n$") %>%
  knitr::kable()
```

## Quick proportions with conficence intervals

There are three functions that will provide quick statistics for different rates
based on binomial estimates of proportions from `binom::binom.wilson()`

 - `attack_rate()`
 - `case_fatality_rate()`
 - `mortality_rate()`

```{r rates}
attack_rate(10, 50)
case_fatality_rate(2, 50)
mortality_rate(40, 50000)
```

In addition, it's possible to rapidly calculate Case fatality rate from a
linelist, stratified by different groups (e.g. gender):

```{r cfr, eval = requireNamespace("outbreaks")}
library("outbreaks")
case_fatality_rate_df(ebola_sim_clean$linelist, 
  outcome == "Death", 
  group = gender,
  add_total = TRUE,
  mergeCI = TRUE
)
```

## Inline functions

The inline functions make it easier to print estimates with confidence intervals
in reports with the correct number of digits.

 - `fmt_ci()` formats confidence intervals from three numbers. (e.g. `fmt_ci(50, 10, 80)` produces `r fmt_ci(50, 10, 80)`
 - `fmt_pci()` formats confidence intervals from three fractions, multiplying by 100 beforehand.

The `_df` suffixes (`fmt_ci_df()`, `fmt_pci_df()`) will print the confidence
intervals for data stored in data frames. These are designed to work with the
outputs of the rates functions. For example, `fmt_ci_df(attack_rate(10, 50))`
will produce `r fmt_ci_df(attack_rate(10, 50))`. All of these suffixes will have
three options `e`, `l`, and `u`. These refer to `estimate`, `lower`, and `upper`
column positions or names. 

 - `fmt_count()` will count a condition in a data frame and present the number
   and percent of `TRUE` values. For example, if you wanted to count the number
   of women patients from Rokupa hospital, you would write: `fmt_count(ebola_sim_clean$linelist, gender == "f", hospital == "Rokupa Hospital")` and it would produce: `r fmt_count(ebola_sim_clean$linelist, gender == "f", hospital == "Rokupa Hospital")`


## Confidence interval manipulation

The confidence interval manipulation functions take in a data frame and combine
their confidence intervals into a single character string much like the inline
functions do. There are two flavors:

 - `merge_ci_df()` and `merge_pci_df()` will merge just the values of the confidence
   interval and leave the estimate alone. Note: this WILL remove the lower and
   upper columns.
 - `unite_ci()` merges both the confidence interval and the estimate into a 
   single character column. This generally has more options than `merge_ci()`

This is useful for reporting models:

```{r unite_ci}
fit <- lm(100/mpg ~ disp + hp + wt + am, data = mtcars)
df  <- data.frame(v = names(coef(fit)), e = coef(fit), confint(fit), row.names = NULL)
names(df) <- c("variable", "estimate", "lower", "upper")
print(df)

# unite CI has more options
unite_ci(df, "slope (CI)", estimate, lower, upper, m100 = FALSE, percent = FALSE)

# merge_ci just needs to know where the estimate is
merge_ci_df(df, e = 2)
```

## Age categories

A couple of functions are dedicated to constructing age categories and partitioning them into separate chunks.

 - `age_categories()` takes in a vector of numbers and returns formatted age categories.
 - `group_age_categories()` will take a data frame with different age
   categories in columns (e.g. years, months, weeks) and combine them into a
   single column, selecting the column with the lowest priority.

```{r ages}
set.seed(1)
x <- sample(0:100, 20, replace = TRUE)
y <- ifelse(x < 2, sample(48, 20, replace = TRUE), NA)
df <- data.frame(
  age_years = age_categories(x, upper = 80), 
  age_months = age_categories(y, upper = 16, by = 6)
)
df %>% 
  group_age_categories(years = age_years, months = age_months)
```

